---
title: "Class 19"
author: 'Shivani Khosla (PID: A59010433)'
date: "12/2/2021"
output:
  pdf_document: default
  html_document: default
---
> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes. Hint: The read.table(), summary() and boxplot() functions will likely be useful here. There is an example R script online to be used ONLY if you are struggling in vein. Note that you can find the medium value from saving the output of the boxplot() function to an R object and examining this object.
There is also the medium() and summary() function that you can use to check your understanding.

reading in file with expression data
```{r}
data <- read.table("rs8067378_ENSG00000172057.6.txt")
```

quantities of each genotype
```{r}
table(data$geno)
```

statistics for A/A genotype: median is 31.25
```{r}
summary(data[data$geno=="A/A",])
```

statistics for A/G genotype: median is 25.065
```{r}
summary(data[data$geno=="A/G",])
```

statistics for G/G genotype: median is 20.074
```{r}
summary(data[data$geno=="G/G",])
```

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3? Hint: An example boxplot is provided overleaf – yours does not need to be as polished as this one.

generate boxplot in base R; by using tilde, expression values are grouped by genotype
```{r}
boxplot(data$exp ~ data$geno)
```

I answered the next part of the question in the lab sheet; I will include the answer here as well: 
The median expression for A/A is higher than for G/G, but there is lots of overlap in their distributions, so the difference may not be statistically significant. I would say that SNP does not affect ORMDL3 expression. 
