---
title: "Class 09 Mini-Project"
author: 'Shivani Khosla (PID: A59010433)'
date: "10/27/2021"
output: github_document
---

read in data
```{r}
fna.data <- "WisconsinCancer.csv"
wisc.df <- read.csv(fna.data, row.names=1)
```

show dataframe
```{r}
#wisc.df
```

remove first column(expert diagonosis)
```{r}
wisc.data <- wisc.df[,-1]
```

expert diagnosis
```{r}
diagnosis <- as.factor(wisc.df[,1])
```

> Q1. How many observations are in this dataset?

```{r}
nrow(wisc.data)
```

> Q2. How many of the observations have a malignant diagnosis?

```{r}
length(which(diagnosis == "M"))
```

> Q3. How many variables/features in the data are suffixed with _mean?

```{r}
sum(grepl("*_mean", colnames(wisc.data)))
```

checking means and standard devs for each column to see if it needs to be scaled
```{r}
colMeans(wisc.data)

apply(wisc.data,2,sd)
```
need to scale because values are vastly different among columns
```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
wisc.pr
```
summary of PCA
```{r}
summary(wisc.pr)
```
> Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

44.27%

> Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

3 PCs, this gives 72.6% of the variance 

> Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

7 PCs, this gives 91% of the variance

```{r}
biplot(wisc.pr)
```
> Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

The column labels are blocking the points, which are listed as numeric values rather than dots. This is very difficult to understand. 

plot for PC1 and PC2
```{r}
plot(wisc.pr$x[,1:2], col = diagnosis , 
     xlab = "PC1", ylab = "PC2")
```

```{r}
plot(wisc.pr$x[,c(1,3)], col = diagnosis, 
     xlab = "PC1", ylab = "PC3")
```

making ggplot scatterplot for PC1 vs PC2 

```{r}
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis

library(ggplot2)

ggplot(df) + 
  aes(PC1, PC2, col=diagnosis) + 
  geom_point()
```

variance of each PC

```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

> Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?

```{r}
wisc.pr$rotation[,1]
```
 -0.26085376

> Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?

```{r}
summary(wisc.pr)
```
5

```{r}
data.scaled <- scale(wisc.data)
```

```{r}
data.dist <- dist(data.scaled)
#data.dist
```

```{r}
wisc.hclust <- hclust(data.dist)
wisc.hclust

```
> Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

h = 19

```{r}
plot(wisc.hclust)
abline(h=19, col="red", lty=2)
```

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, 4)
wisc.hclust.clusters
```
```{r}
table(wisc.hclust.clusters, diagnosis)
```

> Q12. Can you find a better cluster vs diagnoses match by cutting into a different number of clusters between 2 and 10?

Our group determined that a cluster of 4 is optimal, from trying different values. There is good separation between B and M. 

```{r}
wisc.hclust.clusters2 <- cutree(wisc.hclust, 5)
table(wisc.hclust.clusters2, diagnosis)
```
> Q13. Which method gives your favorite results for the same data.dist dataset? Explain your reasoning.

Our group decided that the ward.D2 method splits the data more cleanly than the other methods


```{r}
wisc.hclust2 <- hclust(data.dist, method = "ward.D2")
wisc.hclust2

plot(wisc.hclust2)

```
```{r}
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:4]), method = "ward.D2")
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```

```{r}
table(grps, diagnosis)

```
plot hclust, by group
```{r}
plot(wisc.pr$x[,1:2], col=grps)

```
plot hclust by diagnosis
```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)

```
Find cluster membership vector by cutting tree into k = 2 groups

```{r}
grps <- cutree(wisc.pr.hclust, k = 2)
table(grps)
```

compare to the expert M and B vector

```{r}
table(diagnosis)
```
cross-table by giving the table() function two inputs

```{r}
table(grps, diagnosis)
```

accuracy: how many did we get correct?
```{r}
(165 + 351) / nrow(wisc.data)
```
specificity: tn/(tn + fn)

```{r}
351 / (351 + 47)
```
sensitivity: tp/(tp + fn)

```{r}
165/(165 + 6)
```
predictions
```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
#npc
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

