---
title: "Lab Supplement"
author: 'Shivani Khosla (PID: A59010433)'
date: "10/24/2021"
output:
  pdf_document: default
  html_document: default
---
Original code to make function from: 
```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

To make function, let's make a PDB code the argument/input (pdbCode) for the function. 

This function will output the Bfactor by residue line plot for the particular protein PDB code entered. Specifically, this is done by first excluding all but the 'A' chain of the protein. 

Since in the example, each 'block' of the function is written three times for three different PDB codes, I just took one of each and generalized these lines to 's' rather than 's1', 's2', or 's3'. This is because the new function only takes in one PDB code. 

To call the function, just write improvedPDBFunc() and write the desired PDB code in quotes inside the parenthesis. 
```{r}
improvedPDBFunc <- function(pdbCode) {
  s <- read.pdb(pdbCode)
  s.chainA <- trim.pdb(s, chain="A", elety="CA")
  s.b <- s.chainA$atom$b
  plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor")
}
```

Example function run: 
```{r}
improvedPDBFunc('4AKE')
```

