---
title: "class 17 new"
author: 'Shivani Khosla (PID: A59010433)'
date: "11/28/2021"
output:
  html_document: default
  pdf_document: default
---
```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```
> Q1. What column details the total number of people fully vaccinated? 
  
persons_fully_vaccinated

> Q2. What column details the Zip code tabulation area? 

zip_code_tabulation_area

> Q3. What is the earliest date in this dataset? 

2021-01-05

> Q4. What is the latest date in this dataset? 

2021-11-23

for summary of dataset, use skim function 
```{r}
skimr::skim(vax)
```
> Q5. How many numeric columns are in this dataset? 

9 numeric columns

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```

> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

```{r}
round(sum(is.na(vax$persons_fully_vaccinated)) / length(vax$persons_fully_vaccinated) * 100, 2)
```


> Q8. [Optional]: Why might this data be missing?

As mentioned in the introduction to the assignment, zip codes referring to military bases will not show vaccination data due to the way the database was put together. 


package for working with dates
```{r}
library(lubridate)
```

```{r}
today()
```

```{r}
# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

timeframe of dataset (time between first and last date in this data)
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

> Q9. How many days have passed since the last update of the dataset? 

```{r}
today() - vax$as_of_date[nrow(vax)]
```


> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(table(vax$as_of_date))
```
helpful package to deal with zipcodes
```{r}
library(zipcodeR)
```

centroid of the La Jolla zip code
```{r}
geocode_zip('92037')

```

distance between two zip codes
```{r}
zip_distance('92037','92109')
```

census data for zip codes 
```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
# Subset to San Diego county only areas
sd <- vax[ vax$county == "San Diego" , ]
```

same as above but with dplyr
```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

filtering for data in San Diego county and a population of over 10,000
```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

> Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
length(table(sd.10$zip_code_tabulation_area))
```


> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?


```{r}
sd.10[order(sd.10$age12_plus_population, decreasing = TRUE),][1,'zip_code_tabulation_area']
```

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”? 

```{r}
sd.a <- filter(vax, county == "San Diego")
sd.b <- filter(sd.a, as_of_date == "2021-11-09")
head(sd.b)
```
```{r}
mean(sd.b$percent_of_population_fully_vaccinated, na.rm = TRUE)
```


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
hist(sd.b$percent_of_population_fully_vaccinated)
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
library(ggplot2)

ggplot(ucsd) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```
```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```

> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
ggplot(ucsd) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  geom_hline(yintercept =  mean(vax.36$percent_of_population_fully_vaccinated)) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```

> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram()+
  xlim(0,1)
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
zip92109 <- vax[vax$zip_code_tabulation_area==92109,]
zip92109[zip92109$as_of_date=='2021-11-23',"percent_of_population_fully_vaccinated"] > mean(vax.36$percent_of_population_fully_vaccinated)
```
for 92109: above

```{r}
zip92040 <- vax[vax$zip_code_tabulation_area==92040,]
zip92040[zip92040$as_of_date=='2021-11-23',"percent_of_population_fully_vaccinated"] > mean(vax.36$percent_of_population_fully_vaccinated)
```
for 92040: below

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, vax$age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination rate across California",
       subtitle="Only areas with a population above 36k are shown.") +
  geom_hline(yintercept = mean(vax.36$percent_of_population_fully_vaccinated), linetype = 'dashed')
```

> Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

I personally do not mind meeting for in-person class next week, but I do understand if others feel uncomfortable doing so, especially considering the new CoV-2 variant. Thus, I would be fine with either an in-person or virtual class setting next week. 

