---
title: 'Class 6: R Functions'
author: 'Shivani Khosla (PID: A59010433)'
date: "10/15/2021"
output:
  pdf_document: default
  html_document: default
---

## Quick Rmarkdown intro

We can write text of course just like any file. We can **style text to be bold** or *italic*. 

Do:

- this
- and that
- and another thing

This is more text  
and this is a new line

---

We can include some code: 

```{r}
plot(1:10)
```

```{r}
# This is a comment and will not be passed to R
plot(cars)
```


## Time to write a function 

> Q1. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

Find position of minimum value
```{r}
which.min(student1)
```
Minus sign excludes value from element
Find mean of vector which excludes lowest value
```{r}
student1[-which.min(student1)]
```
```{r}
mean(student1[-which.min(student1)])
```
Does not work for student 2 because which.min(student2) returns 8, not 2
```{r}
mean(student2[-which.min(student2)])

```
```{r}
mean(student2, na.rm = TRUE)
```

```{r}
student3
```

is.na function: TRUE corresponds to the NA values 
```{r}
is.na(student2)
```

Find value of where is.na is TRUE
```{r}
student2[is.na(student2)]
```
Replace NA with 0 and compute mean of the new vector minus the minimum value
```{r}
student.prime <- student2
student.prime[is.na(student.prime)] = 0
mean(student.prime[-which.min(student.prime)])
```
```{r}
student.prime <- student3
student.prime[is.na(student.prime)] = 0
mean(student.prime[-which.min(student.prime)])
```

Make code more simple 

```{r}
x <- student3
x[is.na(x)] = 0
mean(x[-which.min(x)])
```

scenario: entering the data incorrectly (character instead of number)

```{r}
student4 <- c(100, NA, 90, "90", 90, 90, 97, 80)
```

as.numeric changes values into numeric type
```{r}
x <- student4
x <- as.numeric(x)
x[is.na(x)] = 0
mean(x[-which.min(x)])
```
Putting into a function. Components of functions: name, input args, body
```{r}
grade <- function(x) {
  x <- as.numeric(x)
  x[is.na(x)] = 0
  mean(x[-which.min(x)])
}
```

call function - it works
```{r}
grade(student1)
```
## Grade whole class

read notebook
```{r}
gradebook <- "https://tinyurl.com/gradeinput"
scores <- read.csv(gradebook, row.names = 1)
scores
```

apply grade function to each row in scores dataframe
```{r}
ans <- apply(scores, 1, grade)
ans
```
> Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student
overall in the gradebook? [3pts]


```{r}
top_scoring_student <- which.max(ans)
top_scoring_student
```

> Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained
the lowest scores overall? [2pts]

use apply mean function on columns

```{r}
apply(scores, 2, mean)
```

use na.rm = TRUE to get rid of NA values 

```{r}
apply(scores, 2, mean, na.rm = TRUE)

```

but would be better to replace with zeros, then find which hw has the lowest mean
```{r}
mask <- scores
mask[is.na(mask)] = 0
which.min(apply(mask, 2, mean))
```
hw2 has the lowest mean

> Q4. Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

using cor function

```{r}
cor(mask$hw5, ans)
```
use apply to use cor function for every hw

```{r}
apply(mask, 2, cor, ans)
```

hw5 is the best predictor

boxplot

```{r}
boxplot(scores)
```

```{r}
grade2 <- function(x, drop.lowest = TRUE) {
  x <- as.numeric(x)
  x[is.na(x)] = 0
  if(drop.lowest){
    mean(x[-which.min(x)])
  } else {
    mean(x)
  }
}
```

```{r}
grade2(student1, drop.lowest = FALSE)
```

